Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-08-14 paper.tex
Edits on Jonathan's recommendation and added references.

\documentclass[10pt,final,journal]{IEEEtran}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[usenames, dvipsnames]{color}

@@

\author{Mark~H.~Jones, and Prof.~Jonathan~Scott}

@@

    \begin{abstract}
        We investigate the possibility of using streaming cells as a means of harvesting energy from a town water supply.
        We verify streaming cell measurements found in the literature and extend those measurements over a wider range of dimensions.
        We measure the electrical power developed from ten streaming cells over a range of pressure differentials using town water.
        We estimate the amount of energy required to operate a simple data logger and transmitter.
        We estimate the amount of energy available from a typical domestic household based on water usage data.
        Finally, we calculate the dimensions required to build a streaming cell harvester large enough to enable domestic water consumption logging and transmission.
    \end{abstract}

    \section{Introduction}
    Domestic and commercial water metering is becoming increasingly important throughout the world.
    Having a cheap and reliable way to retrieve metering information is an important consideration for utility companies.
    There are two methods available to water utility companies when recording the volume of water delivered to consumers.
    \begin{enumerate}
        \item \emph{Mechanical readout:}
            Involves a meter-reader visually sighting the meter's readout from time to time.
            Generally the reading frequency is kept to a minimum and estimate readings are made in the interim.
        \item \emph{Electronic telemetry:}
            Wireless data transmission is increasingly becoming used as a method of metering data recovery.
            Due to the location of water metering, electrical power must be provided by long life batteries.
            These batteries are usually non-rechargeable and therefore must be replaced once exhausted.
        \end{enumerate}
    {\color{blue}[Suggested adding references to the above items. Not really sure what I would be adding - a specific example would be good]}
    The methods previously mentioned are costly to the utility as they require periodic human intervention.
    Even in the case of wireless data transmission, batteries
    The possibility of harvesting energy at the meter will allow periodic data transmission with less maintenance.
    Removing batteries from electrical meters also reduces unwanted electrical waste.

>>>>

\documentclass[10pt,final,journal]{IEEEtran}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{siunitx}
% \usepackage[dvipsnames]{color}

@@

\author{Mark~H.~Jones, and Jonathan~Scott}

@@

    \begin{abstract}
        We investigate the possibility of using streaming cells as a means of harvesting energy from a town water supply.
        We verify streaming cell measurements found in the literature and extend those measurements over a wider range of dimensions.
        We measure the electrical power developed from ten streaming cells over a range of pressure differentials using town water.
        We estimate the amount of energy required to operate a simple data logger and transmitter.
        We estimate the amount of energy available from a typical domestic household based on water usage data.
        From these estimates we calculate minimum geometric dimensions required to deliver practical power output to a wireless water meter.
        Finally, we comment on the feasibility of using streaming cell technology as a means of harvesting energy from a domestic water supply.
    \end{abstract}

    \section{Introduction}
    Domestic and commercial water metering is becoming increasingly important throughout the world.
    Having a cheap and reliable way to retrieve metering information is an important consideration for utility companies.
    There are two methods available to water utility companies when recording the volume of water delivered to consumers.
    \begin{enumerate}
        \item \emph{Mechanical readout:}
            Involves a meter-reader visually sighting the meter's readout from time to time.
            Generally the reading frequency is kept to a minimum and estimate readings are made in the interim.
        \item \emph{Electronic telemetry:}
            Wireless data transmission is increasingly becoming used as a method of metering data recovery.
            Due to the location of water metering, electrical power is usually provided by long life batteries.
        \end{enumerate}
    Both methods involve costly human intervention on the part of the utility company.
    Wireless telemetry offer savings by removing the need to employ meter readers; however, the batteries used are non-rechargeable and have a maximum life-span of 10 years.\cite{BMeters2014}
    The possibility of harvesting energy at the meter allows frequent data reporting with less maintenance.

    Frequent reporting of water usage enables providers to detect water leaks at consumers' properties.~\cite{Britton2013}
    Removing the batteries from wireless telemetry systems reduces both the total cost of ownership and electrical waste.
